---
title: "Final Project"
author: "Amrita Acharya & Sunni Raleigh"
date: "4/10/2022"
output: html_document
---

```{r include=FALSE}
library(ipumsr)
library(tidyverse)
```

```{r}
# ddi <- read_ipums_ddi("usa_00005.xml")
# data <- read_ipums_micro(ddi)
# 
# data <- data %>%
#   filter(STATEFIP == 8) %>%
#   select(-c(SAMPLE, SERIAL, CBSERIAL))
# 
#  write.csv(data, "colorado_housing_data.csv")
```

## LOAD DATA
```{r}
data <- read.csv("colorado_housing_data.csv") %>%
  mutate(FOODSTMP = ifelse(FOODSTMP == 1, 0, 1))
```


##QUESTION 
What are the most important determinants for someone to be eligible for food stamps in the state of Colorado? 

Which variables will predict the likelihood of someone receiving food stamps or not in Colorado?

Questions for Sanogo:
- We're interested in tree and logit regression for models, what are some other models we could use for classification?  
- What are all the model choosing methods we could use? We were thinking of using CV and lasso. Tuning for trees?
- Feedback on research question.

##PART 1: EXPLORATORY DATA ANALYSIS

```{r}
ggplot(data, aes(x = FOODSTMP)) +
  geom_histogram()
```

Important to notice that most of the people do not receive food stamps.

#### Relationship between rent and receiving foodstamps
```{r}
ggplot(data, aes(y = RENT, x = as.factor(FOODSTMP))) +
  geom_boxplot()
```

```{r}
ggplot(data, aes(y = PERWT, x = as.factor(FOODSTMP))) +
  geom_boxplot()
```
#### Relationship between transit time and receiving foodstamps
```{r}
ggplot(data, aes(y = TRANTIME, x = as.factor(FOODSTMP))) +
  geom_boxplot()
```


##LOGISTIC REGRESSION

```{r}
glm_fit = glm(FOODSTMP~.,
              data = data,
              family = binomial)
summary(glm_fit)
```

```{r}
glm_probs = data.frame(probs = predict(glm_fit, type="response"))
head(glm_probs)
```

```{r}
set.seed(1)

train = data %>%
  dplyr::sample_frac(0.8)

test = data %>%
  anti_join(train)
```

```{r}
glm_2 = glm(FOODSTMP~.,
              data = train, family = binomial)
summary(glm_2)
```

```{r}
glm_probs2 = data.frame(probs = predict(glm_2, newdata = test, type="response"))
head(glm_probs2)
```

```{r}
#create probabilities

glm_pred = glm_probs2 %>%
  mutate(pred = ifelse(probs>.5, "No", "Yes"))

glm_pred = cbind(test, glm_pred)

glm_pred %>% 
  count(pred, FOODSTMP) %>%
  spread(FOODSTMP, n, fill = 0)

glm_pred %>%
  summarize(score = mean(pred == FOODSTMP))
```

##CROSS VALIDATION TO SUBSET PREDICTORS



##PART 2: METHODS 

# Logistic Regression 
We performed a logistic regression 

# Random Forest


