---
title: "Final Project"
author: "Amrita Acharya & Sunni Raleigh"
date: "4/10/2022"
output: html_document
---

## LOAD PACKAGES
```{r include=FALSE}
library(tidyverse)
library(ggplot2)
library(ggcorrplot)

set.seed(1)

# library(randomForest)
```

## LOAD DATA
```{r}
data <- read.csv("colorado_housing_data.csv") %>%
  mutate(FOODSTMP = as.factor(ifelse(FOODSTMP == 1, 1, 0))) %>%
  select(-c(X, YEAR, HHWT, CLUSTER, GQ, STATEFIP, STRATA, PERNUM, PERWT, EMPSTATD, VETSTATD, HCOVANY, HINSIHS))
```

## INTRO 
Food insecurity is a nation wide issue that affects millions of American households every year. The USDA defines it as a lack of consistent access to enough food for an active, healthy life. One of the federal programs implemented in each state, Supplemental Nutrition Assistance Program or SNAP, combats this issue by providing food stamps for those who cannot afford to buy their own food. These stamps can be used in grocery stores on a wide variety of products, including produce, poultry, and dairy. 

This exploration of Colorado comes from the fact that Sunni is from there and Amrita volunteers regularly with food insecure individuals in Northampton. Together, we want to ask, what are the most important determinants for someone to be eligible for food stamps in the state of Colorado? 

This is an important project because food insecurity is linked to poverty, healthcare inaccess and housing insecurity, among other issues. By running these analyses we can start to see the relationships between food insecurity and other problems in the state of Colorado. 

We created our own data extract from IPUMS, a data source that contains census and survey data across the country over various spans of time. We extracted data from the year 2019 in the state of Colorado, and chose various demographic characteristics we were interested in to understand this issue. 

The variables we chose are as follow in this table and key. 

Questions for Sanogo:
- We're interested in tree and logit regression for models, what are some other models we could use for classification?  
- What are all the model choosing methods we could use? We were thinking of using CV and lasso. Tuning for trees?
- Feedback on research question.

### Variable list
- `RENT`: Monthly contract rent  
- `FOODSTMP`: Food stamp recipiency  
- `VALUEH`: House value  
- `RACNUM`: Number of major race groups  
- `HCOVPUB`: Public health insurance coverage   
- `EMPSTAT`: Employment status  
- `OCC`: Occupation  
- `VETSTAT`: Veteran status  
- `TRANWORK`: Means of transportation to work  
- `TRANTIME`: Travel time to work  

### Summary of data
```{r}
summary(data)
```

## PART 1: EXPLORATORY DATA ANALYSIS


### Distribution of response variable
```{r}
count <- data %>%
  group_by(FOODSTMP) %>%
  summarise(total = n())

ggplot(count, aes(x = FOODSTMP, y = total, fill = FOODSTMP)) +
  geom_col(show.legend = FALSE) +
  geom_label(aes(label = total), show.legend = FALSE)
  
```
Figure 1. The majority of people in this dataset (53319) received foodstamps in 2019 whereas only 3823 people do not receive foodstamps in that same year. 

Important to notice that most of the people receive food stamps.

#### Relationship between rent and receiving foodstamps
```{r}
ggplot(data, aes(y = RENT, x = FOODSTMP)) +
  geom_boxplot()
```
Figure 2. Individuals who receive foodstamps tend to have a lower average rent than the individuals who do not receive food stamps. Additionally, the spread in the rent values  is much greater for those with food stamps, than without. 

#### Relationship between transit time and receiving foodstamps
```{r}
ggplot(data, aes(y = TRANTIME, x = FOODSTMP)) +
  geom_boxplot()
```
Figure 3. Those receiving food stamps have a higher transit time on average. However, the spread of the data is similar between both categories. 

#### Correlation of variables
```{r}
model.matrix(~0+., data = data) %>%
  cor(use = "pairwise.complete.obs") %>%
  ggcorrplot(lab = TRUE, lab_size = 2)
```
Figure 4. Rent and house value are highly correlated at a value of 0.79. The higher the house value the higher the rent is. Neither food stamp variables are highly correlated with the other variables. The highest correlation value is between Food stamps and HCOVANY which is whether or not someone has public health insurance. This is at a value of 0.25. 

## PART 2: METHODOLOGY
We split the data into a training set and a testing set.

```{r}
train <- data %>%
  dplyr::sample_frac(0.8)

test <- data %>%
  anti_join(train)
```


### MODEL 1: LOGISTIC REGRESSION

We performed a logistic regression on our data first to understand the probabilities of someone getting food stamps in Colorado. We wanted to use logistic regression first because it is easy to interpret for the sake of this research project, and it is a model we are very familiar with. Our food stamp variable FOODSTMP has a binary response, either 0, not receiving food stamps, or 1, receiving food stamps. Becaues of this, we decided on logistic regression because it works with binary responses. Logistic regression also has low variance because it assumes a linear relationship between the predictor and resposne.

#### Model for entire dataset

```{r}
glm_fit <- glm(FOODSTMP~.,
              data = data,
              family = binomial)
summary(glm_fit)
```

```{r}
glm_probs <- data.frame(probs = predict(glm_fit, type="response"))
head(glm_probs)
```

#### Training model

```{r}
glm_2 <- glm(FOODSTMP~., data = train, family = binomial)
summary(glm_2)
```

```{r}
glm_probs2 <- data.frame(probs = predict(glm_2, newdata = test, type="response"))
head(glm_probs2)
```

```{r}
#create probabilities

glm_pred <- glm_probs2 %>%
  mutate(pred = ifelse(probs>.5, "No", "Yes"))

glm_pred <- cbind(test, glm_pred)

glm_pred %>% 
  count(pred, FOODSTMP) %>%
  spread(FOODSTMP, n, fill = 0)

glm_pred %>%
  summarize(score = mean(pred == FOODSTMP))
```

### MODEL 2: BAGGING

Based on recommendations from Professor Sanogo, we decided to also use tree based methods for its interpretability and because our dataset has many predictors. We decided on bagging because it reduces variance in our data. 

Is the computational power worth it?? 


#### Bagging Model 
https://discuss.analyticsvidhya.com/t/what-does-the-warning-the-response-has-five-or-fewer-unique-values-while-building-random-forest-mean/6442

```{r}
# bagging_model <- randomForest(FOODSTMP~.,
#                          data = train,
#                          mtry = 9,
#                          importance = TRUE
#                          )
# 
# saveRDS(bagging_model, "bag_mod.rds")

bagging_model <- readRDS("bag_mod.rds")

summary(bagging_model)
```

```{r}
bagging_estimate <- predict(bagging_model, newdata = test)

mean(test$FOODSTMP == bagging_estimate) # accuracy rate

table(test$FOODSTMP, bagging_estimate) # confusion matrix
```

```{r}
importance(bagging_model)

varImpPlot(bagging_model)
```


## PART 3: KEY TAKEAWAYS 

EDA Takeaways: 
- Healthcare foodstamp correlation/further questions? --> are people even receiving health insurance?
- Rent/ food stamp --> section 8?
- 

